---
title: "Assignment 6: FDR"
author: "Anthony Li and Naman Shah"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up required packages

```{r}
source("regday3.pck")
source("regday3.pck")
source("fdr.pck")
library("ISLR")
library("infer")
library("moderndive")
```

Let's have a look at the data.
```{r}
attributes(NCI60)
summary(NCI60)
unique(NCI60$labs)
```

Let's save the dataset as "cancer".
```{r}
cancer <- NCI60
```

# Work with CNS, Melanoma, Leukemia
## FDR = 0.2
Let's take an FDR of 0.2

First let's subset the data into cns, mela, and leuk.
```{r}
cns <- NCI60$data[NCI60$labs == 'CNS',]
mela <- NCI60$data[NCI60$labs == 'MELANOMA',]
leuk <- NCI60$data[NCI60$labs == 'LEUKEMIA',]
```
Let's apply FDR.
```{r}
cnsData <- fdr(apply(cns, 2, myt), 0.2, ind = F)
```

No relevant data for CNS.
Let's make melaData and leukData.
```{r}
melaData <- fdr(apply(mela, 2, myt), 0.2, ind = F)

leukData <- fdr(apply(leuk, 2, myt), 0.2, ind = F)
```


Let's look at the interesting and make an interecstion.
```{r}
leukV <- leukData$interesting

melaV <- melaData$interesting

commonGene <- intersect(melaV, leukV)
commonGene
```

The commonGene has nothing, meaning there are no genes in common.
Let's try the same approach for a higher FDR, like 0.3.

# FDR - 0.3
```{r}
cnsData <- fdr(apply(cns, 2, myt), 0.3, ind = F)

melaData <- fdr(apply(mela, 2, myt), 0.3, ind = F)

leukData <- fdr(apply(leuk, 2, myt), 0.3, ind = F)

cnsV <- cnsData$interesting

leukV <- leukData$interesting

melaV <- melaData$interesting

commonGene01 <- intersect(cnsV, melaV)
print("Common genes between cns and mela.")
commonGene01
length(commonGene01)

commonGene02 <- intersect(melaV, leukV)
print("Common genes between mela and leuk")
commonGene02
length(commonGene02)

commonGene03 <- intersect(leukV, cnsV)
print("Common genes between leuk and cns")
commonGene03
length(commonGene03)

commonGene1 <- intersect(commonGene01, commonGene02)
print("Common genes between all the three datasets")
commonGene1
length(commonGene1)

```

The gene is 4289 is common between mela and leuk for the larger FDR value of 0.3


# FDR - 0.4
```{r}
cnsData <- fdr(apply(cns, 2, myt), 0.4, ind = F)

melaData <- fdr(apply(mela, 2, myt), 0.4, ind = F)

leukData <- fdr(apply(leuk, 2, myt), 0.4, ind = F)

cnsV <- cnsData$interesting

leukV <- leukData$interesting

melaV <- melaData$interesting

commonGene01 <- intersect(cnsV, melaV)
print("Common genes between cns and mela.")
commonGene01
length(commonGene01)

commonGene02 <- intersect(melaV, leukV)
print("Common genes between mela and leuk")
commonGene02
length(commonGene02)

commonGene03 <- intersect(leukV, cnsV)
print("Common genes between leuk and cns")
commonGene03
length(commonGene03)

commonGene1 <- intersect(commonGene01, commonGene02)
print("Common genes between all the three datasets")
commonGene1
length(commonGene1)


```

We again get 4289 between mela and leuk.


# FDR - 0.45
```{r}
cnsData <- fdr(apply(cns, 2, myt), 0.45, ind = F)

melaData <- fdr(apply(mela, 2, myt), 0.45, ind = F)

leukData <- fdr(apply(leuk, 2, myt), 0.45, ind = F)

cnsV <- cnsData$interesting

leukV <- leukData$interesting

melaV <- melaData$interesting

commonGene01 <- intersect(cnsV, melaV)
print("Common genes between cns and mela.")
commonGene01
length(commonGene01)

commonGene02 <- intersect(melaV, leukV)
print("Common genes between mela and leuk")
commonGene02
length(commonGene02)

commonGene03 <- intersect(leukV, cnsV)
print("Common genes between leuk and cns")
commonGene03
length(commonGene03)

commonGene1 <- intersect(commonGene01, commonGene02)
print("Common genes between all the three datasets")
commonGene1
length(commonGene1)

```

We get our largest intersection yet between mela and leuk, in six genes. We get 4289, 6369, 4287, 5887, 6080, and 5931 as  common genes  between the two diseases.
Let's have our final try.

# FDR - 0.5
We cannot use any inference from an estimate for 50%, as that is highly likely to be random. This experiment is only for our academic purposes.
```{r}
cnsData <- fdr(apply(cns, 2, myt), 0.5, ind = F)

melaData <- fdr(apply(mela, 2, myt), 0.5, ind = F)

leukData <- fdr(apply(leuk, 2, myt), 0.5, ind = F)

cnsV <- cnsData$interesting

leukV <- leukData$interesting

melaV <- melaData$interesting

commonGene01 <- intersect(cnsV, melaV)
print("Common genes between cns and mela.")
commonGene01
length(commonGene01)

commonGene02 <- intersect(melaV, leukV)
print("Common genes between mela and leuk")
commonGene02
length(commonGene02)

commonGene03 <- intersect(leukV, cnsV)
print("Common genes between leuk and cns")
commonGene03
length(commonGene03)

commonGene1 <- intersect(commonGene01, commonGene02)
print("Common genes between all the three datasets")
commonGene1
length(commonGene1)


```
We have 9 common genes between mela and leuk. We get 4288, 4289, 6369, 4287, 5887, 6080, 5931, 5356, and 6081, as common genes  between the two diseases.

# Conclusions
We tested our data on various percentages, 20%, 30%, 40% and 45%. While we did not get any common genes for 20%, we received 4289 as the only common gene in 30% and 40%. At a very high percentage like 45%, we recived 6 common genes, but due to the high percentage. We cannot determine much from it.
It is important to note that this came from two cancers, melanoma and leukemia, and the gene 4289 might not be statistically independent for this particular gene.
This tells us that the genes from the same cell might not be statistically independent. Mainly, we feel like the gene 4289 might be common in leukemia and melanoma. 


