---
title: "Assignment 01: Regression"
author: "Naman Shah"
date: "`r Sys.Date()`"
output: pdf_document
---

## Assignment 1 - Getting started with Regression

Let's load the data and the libraries


```{r, results='hide', include=FALSE, echo=TRUE}
library("tidyverse")
library("moderndive")
library("skimr")
source("regday1.pck")
```
```{r}
regday1.pck
```

Let's explore the data and the functions


```{r}
head(NOAAGISSWD)
summary(NOAAGISSWD)

gauss.reg1 <-
function(x,y,lambda=0.06,xcol=4,do.plot=T)
{
o1<-order(x)
x1<-x[o1]
y1<-y[o1]
r1<-range(x)
smat<-NULL
n1<-length(x1)
for(i in 1:n1){
        v1<-dnorm(x1,x1[i],lambda)
        v1<-v1/sum(v1)
        H1<-my.hat.w(x1,v1)
        smat<-rbind(smat,H1[i,])
}
yhat<-smat%*%y1
if(do.plot){
lines(x1,yhat,col=xcol)
}
n99<-length(x1)
dferror<-length(x1)-sum(diag(2*smat-smat%*%(t(smat))))
delta1<-sum(diag(t(diag(n99)-smat)%*%(diag(n99)-smat)))
R<-t(diag(n99)-smat)%*%(diag(n99)-smat)
delta2<-2*sum(diag(R%*%R))
resid<-y1-smat%*%y1
ypred<-y1
ypred[o1]<-smat%*%y1
PRESS<-sum((resid/(1-diag(smat)))^2)
list(smat=smat,df=sum(diag(smat)),dferror=dferror,delta1=delta1,delta2=delta2,resid=resid,pred.out=ypred[o1],x.out=x1,press=PRESS,cor=cor(ypred,y))
}

```

We're setting lamda to 0.06 to work with temperature.
For the first set of regression graphs, we will take x to be temperature and assign the y values to the following:

* Drought Count

* Flooding Count

* Severe Storm Count

* Tropical Cyclone Count

* Wildfire Count

* Winter Storm Count

* All Disasters Count

Let's start with **Drought Count** compared with **Delta Temperature**
```{r}
unique(NOAAGISSWD$Drought.Count)
summary(NOAAGISSWD$delta.temp)

plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Drought.Count, main="Correlation between delta temperature and drought counts", xlab="Temperature", ylab="Drought Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Drought.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Drought.Count)
```
We notice that the unique values for drought count are only 0 and 1.
This relationship is a non functional relationship, so a linear regression model is not recommended.
For the bootmansmooth model, we don't see a bar come up, meaning the relationship between delta temperature and drought count is non functional.

Next, **Flooding Count** compared with **Delta Temperature**
```{r}
unique(NOAAGISSWD$Flooding.Count)
plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Flooding.Count, main="Correlation between delta temperature and flood counts", xlab="Temperature", ylab="Flood Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Flooding.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Flooding.Count)
```
The bar at bootmansmooth density graph leans towards the left, indicating the model might not be the best way to model a relationship between flood count and delta temperature.

Next, **Severe Storm Count** compared with **Delta Temperature**
```{r}
unique(NOAAGISSWD$Severe.Storm.Count)
summary(NOAAGISSWD$Severe.Storm.Count)

plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Severe.Storm.Count, main="Correlation between delta temperature and severe storm counts", xlab="Temperature", ylab="Severe Storm Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Severe.Storm.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Severe.Storm.Count)
```
The severe storm counts range from 0 to 19, with some outliers along the way. There is definitely an increase in severe storm counts with the rise in delta temperature.
Again, the bar at bootmansmooth density graph leans towards the left, indicating the model might not be the best way to model a relationship between severe storm count and delta temperature.
The bar plot does seem to indicate that the higher delta temperature is somewhat related to higher severe storm counts.

Next, **Tropical Cyclone Count** compared with **Delta Temperature**
```{r}
unique(NOAAGISSWD$Tropical.Cyclone.Count)
plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Tropical.Cyclone.Count, main="Correlation between delta temperature and tropical cyclone counts", xlab="Temperature", ylab="Tropical Cyclone Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Tropical.Cyclone.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Tropical.Cyclone.Count)
```
Once again, the bar at bootmansmooth density graph leans way towards the left, indicating the model might not be the best way to model a relationship between tropical cyclone count and delta temperature.

Next, **Wildfire Count** compared with **Delta Temperature**
```{r}
unique(NOAAGISSWD$Wildfire.Count)
plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Wildfire.Count, main="Correlation between delta temperature and wildfire counts", xlab="Temperature", ylab="Wildfire Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Wildfire.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Wildfire.Count)
```
We notice that Wildfire Counts can be represented as 0 and 1
This itself proves to be not the best relationship to model as a functional relationship
As indicated by the third graph on the density graph of the smooth model, the bar tends to be more to the left.
This means this relationship isn't functional.

Next, **Winter Storm Count** compared with **Delta Temperature**
```{r}
unique(NOAAGISSWD$Winter.Storm.Count)
plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Winter.Storm.Count, main="Correlation between delta temperature and winter storm counts", xlab="Temperature", ylab="Winter Storm Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Winter.Storm.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Winter.Storm.Count)
```
The data is evenly spread between 0 1 and 2 for delta temperature.
Once again, the bar doesn't show up, telling us this isn't best way to represent a functional relationship.
Linear regression would not be my advice for this model.

Finally, **All Disasters Count** compared with **Delta Temperature**
```{r}
summary(NOAAGISSWD$All.Disasters.Count)
unique(NOAAGISSWD$All.Disasters.Count)

plot(NOAAGISSWD$delta.temp, NOAAGISSWD$All.Disasters.Count, main="Correlation between delta temperature and all disasters counts", xlab="Temperature", ylab="Disaster Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$All.Disasters.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$All.Disasters.Count)
```
All disaster counts defintely shows something of a linear relationship with delta temperature. 
Despite this, the density graphs tells us that the bar is leaning way towards the left, indicating that linear regression isn't the best way to do this.
