---
title: 'Assignment 01: Regression'
author: "Naman Shah"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

## Assignment 1 - Getting started with Regression

Let's load the data and the libraries

```{r, results='hide', include=FALSE, echo=TRUE}
library("tidyverse")
library("moderndive")
library("skimr")
source("regday1.pck")
```

```{r}
regday1.pck
```

Let's explore the data and the functions

```{r}
# Displaying the first few rows of the data
head(NOAAGISSWD)

# Summary of the data in the dataframe
summary(NOAAGISSWD)

# Editing functions to include lambda to be 0.06
gauss.reg1 <-
function(x,y,lambda=0.06,xcol=4,do.plot=T)
{
o1<-order(x)
x1<-x[o1]
y1<-y[o1]
r1<-range(x)
smat<-NULL
n1<-length(x1)
for(i in 1:n1){
        v1<-dnorm(x1,x1[i],lambda)
        v1<-v1/sum(v1)
        H1<-my.hat.w(x1,v1)
        smat<-rbind(smat,H1[i,])
}
yhat<-smat%*%y1
if(do.plot){
lines(x1,yhat,col=xcol)
}
n99<-length(x1)
dferror<-length(x1)-sum(diag(2*smat-smat%*%(t(smat))))
delta1<-sum(diag(t(diag(n99)-smat)%*%(diag(n99)-smat)))
R<-t(diag(n99)-smat)%*%(diag(n99)-smat)
delta2<-2*sum(diag(R%*%R))
resid<-y1-smat%*%y1
ypred<-y1
ypred[o1]<-smat%*%y1
PRESS<-sum((resid/(1-diag(smat)))^2)
list(smat=smat,df=sum(diag(smat)),dferror=dferror,delta1=delta1,delta2=delta2,resid=resid,pred.out=ypred[o1],x.out=x1,press=PRESS,cor=cor(ypred,y))
}

```

We're setting lamda to 0.06 to work with temperature. For the first set of regression graphs, we will take x to be temperature and assign the y values to the following:

-   Drought Count

-   Flooding Count

-   Severe Storm Count

-   Tropical Cyclone Count

-   Wildfire Count

-   Winter Storm Count

-   All Disasters Count

### Let's start with **Drought Count** compared with **Delta Temperature**

```{r}
unique(NOAAGISSWD$Drought.Count)
summary(NOAAGISSWD$delta.temp)

plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Drought.Count, main="Correlation between delta temperature and drought counts", xlab="Temperature", ylab="Drought Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Drought.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Drought.Count)
```

We notice that the unique values for drought count are only 0 and 1. This relationship is a non functional relationship, so a linear regression model is not recommended. For the bootmansmooth model, we don't see a bar come up, meaning the relationship between delta temperature and drought count is non functional. As per the first graph, I would also say that the relationship is non functional.

### Next, **Flooding Count** compared with **Delta Temperature**

```{r}
unique(NOAAGISSWD$Flooding.Count)
plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Flooding.Count, main="Correlation between delta temperature and flood counts", xlab="Temperature", ylab="Flood Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Flooding.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Flooding.Count)
```

The bar at bootmansmooth density graph leans towards the left, indicating the model might not be the best way to model a relationship between flood count and delta temperature. I would say that there seems to be slighlt linear relationship between delta temps 0.8 to 1.0, but overall the relationship seems to be non-linear. I would not recommend the bootmansmooth model despite indications that relationship might be functional, because there is simply a lot of disparity.

### Next, **Severe Storm Count** compared with **Delta Temperature**

```{r}
unique(NOAAGISSWD$Severe.Storm.Count)
summary(NOAAGISSWD$Severe.Storm.Count)

plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Severe.Storm.Count, main="Correlation between delta temperature and severe storm counts", xlab="Temperature", ylab="Severe Storm Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Severe.Storm.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Severe.Storm.Count)
```

The severe storm counts range from 0 to 19, with some outliers along the way. There is definitely an increase in severe storm counts with the rise in delta temperature. Again, the bar at bootmansmooth density graph leans towards the left, indicating the linear model might not be the best way to model a relationship between severe storm count and delta temperature. The bar plot does seem to indicate that the higher delta temperature is somewhat related to higher severe storm counts.

I would say the relationship is functional, but not linear. It seems to be some sort of an exponential function, because with higher delta temperatures, the values of severe storm counts go up by huge increase.

### Next, **Tropical Cyclone Count** compared with **Delta Temperature**

```{r}
unique(NOAAGISSWD$Tropical.Cyclone.Count)
summary(NOAAGISSWD$Tropical.Cyclone.Count)
summary(NOAAGISSWD$delta.temp)

NOAAGISSWD[NOAAGISSWD$Tropical.Cyclone.Count==7, ]

plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Tropical.Cyclone.Count, main="Correlation between delta temperature and tropical cyclone counts", xlab="Temperature", ylab="Tropical Cyclone Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Tropical.Cyclone.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Tropical.Cyclone.Count)
```

Once again, the bar at bootmansmooth density graph leans way towards the left, indicating the linear model might not be the best way to model a relationship between tropical cyclone count and delta temperature.

I would say this is a non functional relationship, simply because of the outliers and the inconsistencies in data for the increase in temeperature. At 1.07 delta temperature in 2020, there were 7 tropical cyclones, compared

### Next, **Wildfire Count** compared with **Delta Temperature**

```{r}
unique(NOAAGISSWD$Wildfire.Count)
plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Wildfire.Count, main="Correlation between delta temperature and wildfire counts", xlab="Temperature", ylab="Wildfire Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Wildfire.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Wildfire.Count)
```

We notice that Wildfire Counts can be represented as 0 and 1 This itself proves to be not the best relationship to model as a functional relationship As indicated by the third graph on the density graph of the smooth model, the bar tends to be more to the left. This means this relationship isn't functional.

With a function that can output 0 and 1, in a much random fashion, but we can see that with bigger delta temperature, the wildfire counts do tend to be 1.

### Next, **Winter Storm Count** compared with **Delta Temperature**

```{r}
unique(NOAAGISSWD$Winter.Storm.Count)
plot(NOAAGISSWD$delta.temp, NOAAGISSWD$Winter.Storm.Count, main="Correlation between delta temperature and winter storm counts", xlab="Temperature", ylab="Winter Storm Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$Winter.Storm.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$Winter.Storm.Count)
```

The data is evenly spread between 0 1 and 2 for delta temperature. Once again, the bar doesn't show up, telling us this isn't best way to represent a functional relationship. Linear regression would not be my advice for this model.

The bar shows up around 0.0 for linear model, while the density lies between 0.3 to 0.9. So a linear relationship isn't recommended. Even for the smooth function, the data isn't functional, so a functional relationship is also not the best option.

### Finally, **All Disasters Count** compared with **Delta Temperature**

```{r}
summary(NOAAGISSWD$All.Disasters.Count)
unique(NOAAGISSWD$All.Disasters.Count)

plot(NOAAGISSWD$delta.temp, NOAAGISSWD$All.Disasters.Count, main="Correlation between delta temperature and all disasters counts", xlab="Temperature", ylab="Disaster Count", col="red")

bootmatsmooth(NOAAGISSWD$delta.temp, NOAAGISSWD$All.Disasters.Count)
bootmatlin(NOAAGISSWD$delta.temp, NOAAGISSWD$All.Disasters.Count)
```

Personally, this would have been my best guess for a linear model. But unfortunately, as indicated by the density graphs, this isn't the best fit for a linear model. This however, is an increasing functional relationship, and my bet would be on an exponential model as the values seem to increase by a notch as the temperature grows.
