---
title: 'Assignment 4'
author: "Naman Shah"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

# Loading the libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("ISLR")
source("regboot.pck")
source("fixregday1.pck")

weather <- NOAAGISSWD
default <- Default
```
# Running lmboot on Severe Storms

## Severe Storms against Delta Temperature
```{r}
lmboot(Severe.Storm.Count~delta.temp, weather, 5000)

```
No zeroes for Severe Storms against Delta Temperature
Press - 308.5
Confidence - 95%
## Severe Storms against Year

```{r}
lmboot(Severe.Storm.Count~Year, weather, 5000)
```

No zeroes for Severe Storms against Year
Press - 278.5
Confidence - 95%

## Severe Storms against (Year + Temperature)

```{r}
lmboot(Severe.Storm.Count~I(Year + delta.temp), weather, 5000)
```

No zeroes for Severe Storms against (Year + Temperature)
Press - 277.4
Confidence - 95%

## Severe Storms against (Year x Temperature)

```{r}
lmboot(Severe.Storm.Count~I(Year * delta.temp), weather, 5000)
```

No zeroes for Severe Storms against (Year x Temperature)
Press - 304.3
Confidence - 95%

## Severe Storms against (Year^2)

```{r}
lmboot(Severe.Storm.Count~I(Year^2), weather, 5000)
```

No zeroes for Severe Storms against (Year^2)
Press - 276.8
Confidence - 95%

## Severe Storms against (Temperature^2)

```{r}
lmboot(Severe.Storm.Count~I(delta.temp^2), weather, 5000)
```

There is a zero for Severe Storms against (Temperature^2)
Press - 277.5
Confidence - 95%

Let's try to add a one

## Severe Storms against (Temperature^2)+1

```{r}
lmboot(Severe.Storm.Count~(I(delta.temp^2) + 1), weather)
```

There is still a zero for Severe Storms against (Temperature^2)+1
Press - 277.5
Confidence - 95%

This is not a feasible model.

## Severe Storms against (Year + Temperature + (Year*Temperature))

```{r}
lmboot(Severe.Storm.Count~I(Year + delta.temp + (Year*delta.temp)), weather, 5000)
```

No zeroes for Severe Storms against (Year + Temperature + (Year*Temperature))
Press - 301.8
Confidence - 95%


## Severe Storms against (Year + Temperature + (Year^2) + (Temerature^2))

```{r}
lmboot(Severe.Storm.Count~Year + delta.temp + I(Year^2) + I(delta.temp^2), weather, 5000)
```

There are zeroes for Severe Storms against (Year + Temperature + (Year^2) + (Temerature^2))
Press - 154.8
Confidence - 95%

Let's try removing a one.


## Severe Storms against (Year + Temperature + (Year^2) + (Temerature^2)) - 1

```{r}
lmboot(Severe.Storm.Count~Year + delta.temp + I(Year^2) + I(delta.temp^2) - 1, weather, 5000)
```

No zeroes for Severe Storms against (Year + Temperature + (Year^2) + (Temerature^2)) - 1
Press - 181.6
Confidence - 95%


## Severe Storms against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temerature^2))

```{r}
lmboot(Severe.Storm.Count~Year + delta.temp + I(delta.temp*Year) + I(Year^2) + I(delta.temp^2), weather, 5000)
```

There are zeroes for Severe Storms against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temerature^2))
Press - 143.6
Confidence - 95%

Let's try removing a one


## Severe Storms against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temperature^2)) - 1

```{r}
lmboot(Severe.Storm.Count~Year + delta.temp + I(delta.temp*Year) + I(Year^2) + I(delta.temp^2) - 1, weather, 5000)
```

There are still zeroes for Severe Storms against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temerature^2))
Press - 185.6
Confidence - 95%

Let's try adding a one


## Severe Storms against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temperature^2)) + 1

```{r}
lmboot(Severe.Storm.Count~Year + delta.temp + I(delta.temp*Year) + I(Year^2) + I(delta.temp^2) + 1, weather, 5000)
```

There are still some zeroes for Severe Storms against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temperature^2)) + 1
Press - 143.6
Confidence - 95%

These models are not feasible.


# Running lmboot on All Disasters

## All Disasters against Delta Temperature
```{r}
lmboot(All.Disasters.Count ~ delta.temp, weather, 5000)
```
No zeroes for All Disasters against Delta Temperature
Press - 651.2
Confidence - 95%

## All Disasters against Year
```{r}
lmboot(All.Disasters.Count ~ Year, weather, 5000)
```
No zeroes for All Disasters against Year
Press - 597.1
Confidence - 95%

## All Disasters against (Year + Temperature)
```{r}
lmboot(All.Disasters.Count ~ I(Year + delta.temp), weather, 5000)
```
No zeroes for All Disasters against (Year + Temperature)
Press - 594.8
Confidence - 95%

## All Disasters against (Year x Temperature)
```{r}
lmboot(All.Disasters.Count ~ I(Year * delta.temp), weather, 5000)
```
No zeroes for All Disasters against (Year x Temperature)
Press - 643.3
Confidence - 95%

## All Disasters against (Year^2)
```{r}
lmboot(All.Disasters.Count ~ I(Year^2), weather, 5000)
```
No zeroes for All Disasters against (Year^2)
Press - 594.2
Confidence - 95%

## All Disasters against (Temperature^2)
```{r}
lmboot(All.Disasters.Count ~ I(delta.temp^2), weather, 5000)
```
No zeroes for All Disasters against (Temperature^2)
Press - 492.6
Confidence - 95%

## All Disasters against (Year + Temperature + (Year*Temperature))
```{r}
lmboot(All.Disasters.Count ~ I(Year + delta.temp + (Year*delta.temp)), weather, 5000)
```
No zeroes for All Disasters against (Year + Temperature + (Year*Temperature))
Press - 638.5
Confidence - 95%

## All Disasters against (Year + Temperature + (Year^2) + (Temerature^2))
```{r}
lmboot(All.Disasters.Count ~ I(Year + delta.temp + (Year^2) + (delta.temp^2)), weather, 5000)
```
No zeroes for All Disasters against (Year + Temperature + (Year^2) + (Temerature^2))
Press - 594.2
Confidence - 95%

## All Disasters against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temerature^2))
```{r}
lmboot(All.Disasters.Count ~ I(Year + delta.temp + (delta.temp*Year) + (Year^2) + (delta.temp^2)), weather, 5000)
```
No zeroes for All Disasters against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temerature^2))
Press - 592.9
Confidence - 95%

#Conclusions
For both the counts, Severe Storm and All Disaster, any combination of Year and Temperature seems to do the job. None of the lmboots returned any zeroes. In this case, the lowest press becomes the decider of the function.
For severe storms, the lowest press is 276.2 for Severe Storms against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temperature^2))
For all disasters, the lowest press is 492.6 for All Disasters against (Temperature^2).
What is interesting about all disasters is that Temperature squared has a significantly lower values than other models, showing an increased effectiveness. On the other hand, a number of models for severe storms are in the 270-310 range.

