---
title: 'Assignment 4'
author: "Naman Shah"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

# Loading the libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("ISLR")
source("regboot.pck")
source("fixregday1.pck")

weather <- NOAAGISSWD
default <- Default
```
# Running lmboot on Severe Storms
## Severe Storms against Delta Temperature
```{r}
lmboot(Severe.Storm.Count~delta.temp, weather, 5000)

```
No zeroes for Severe Storms against Delta Temperature
Press - 308.5
Confidence - 95%
## Severe Storms against Year

```{r}
lmboot(Severe.Storm.Count~Year, weather, 5000)
```

No zeroes for Severe Storms against Year
Press - 278.5
Confidence - 95%

## Severe Storms against (Year + Temperature)

```{r}
lmboot(Severe.Storm.Count~I(Year + delta.temp), weather, 5000)
```

No zeroes for Severe Storms against (Year + Temperature)
Press - 277.4
Confidence - 95%

## Severe Storms against (Year x Temperature)

```{r}
lmboot(Severe.Storm.Count~I(Year * delta.temp), weather, 5000)
```

No zeroes for Severe Storms against (Year x Temperature)
Press - 304.3
Confidence - 95%

## Severe Storms against (Year^2)

```{r}
lmboot(Severe.Storm.Count~I(Year^2), weather, 5000)
```

No zeroes for Severe Storms against (Year^2)
Press - 276.8
Confidence - 95%

## Severe Storms against (Temperature^2)

```{r}
lmboot(Severe.Storm.Count~I(delta.temp^2), weather, 5000)
```

No zeroes for Severe Storms against (Temperature^2)
Press - 277.5
Confidence - 95%

## Severe Storms against (Year + Temperature + (Year*Temperature))

```{r}
lmboot(Severe.Storm.Count~I(Year + delta.temp + (Year*delta.temp)), weather, 5000)
```

No zeroes for Severe Storms against (Year + Temperature + (Year*Temperature))
Press - 301.8
Confidence - 95%

## Severe Storms against (Year + Temperature + (Year^2) + (Temerature^2))

```{r}
lmboot(Severe.Storm.Count~I(Year + delta.temp + (Year^2) + (delta.temp^2)), weather, 5000)
```

No zeroes for Severe Storms against (Year + Temperature + (Year^2) + (Temerature^2))
Press - 276.8
Confidence - 95%

## Severe Storms against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temerature^2))

```{r}
lmboot(Severe.Storm.Count~I(Year + delta.temp + (delta.temp*Year) + (Year^2) + (delta.temp^2)), weather, 5000)
```

No zeroes for Severe Storms against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temerature^2))
Press - 276.2
Confidence - 95%
Sure, here's the modified R Markdown code with "Severe Storms" replaced with "All Disasters" and "Severe.Storm.Count" replaced with "All.Disaster.Count":

#Running lmboot on All Disasters
## All Disasters against Delta Temperature
```{r}
lmboot(All.Disasters.Count ~ delta.temp, weather, 5000)
```
No zeroes for All Disasters against Delta Temperature
Press - 651.2
Confidence - 95%

## All Disasters against Year
```{r}
lmboot(All.Disasters.Count ~ Year, weather, 5000)
```
No zeroes for All Disasters against Year
Press - 597.1
Confidence - 95%

## All Disasters against (Year + Temperature)
```{r}
lmboot(All.Disasters.Count ~ I(Year + delta.temp), weather, 5000)
```
No zeroes for All Disasters against (Year + Temperature)
Press - 594.8
Confidence - 95%

## All Disasters against (Year x Temperature)
```{r}
lmboot(All.Disasters.Count ~ I(Year * delta.temp), weather, 5000)
```
No zeroes for All Disasters against (Year x Temperature)
Press - 643.3
Confidence - 95%

## All Disasters against (Year^2)
```{r}
lmboot(All.Disasters.Count ~ I(Year^2), weather, 5000)
```
No zeroes for All Disasters against (Year^2)
Press - 594.2
Confidence - 95%

## All Disasters against (Temperature^2)
```{r}
lmboot(All.Disasters.Count ~ I(delta.temp^2), weather, 5000)
```
No zeroes for All Disasters against (Temperature^2)
Press - 492.6
Confidence - 95%

## All Disasters against (Year + Temperature + (Year*Temperature))
```{r}
lmboot(All.Disasters.Count ~ I(Year + delta.temp + (Year*delta.temp)), weather, 5000)
```
No zeroes for All Disasters against (Year + Temperature + (Year*Temperature))
Press - 638.5
Confidence - 95%

## All Disasters against (Year + Temperature + (Year^2) + (Temerature^2))
```{r}
lmboot(All.Disasters.Count ~ I(Year + delta.temp + (Year^2) + (delta.temp^2)), weather, 5000)
```
No zeroes for All Disasters against (Year + Temperature + (Year^2) + (Temerature^2))
Press - 594.2
Confidence - 95%

## All Disasters against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temerature^2))
```{r}
lmboot(All.Disasters.Count ~ I(Year + delta.temp + (delta.temp*Year) + (Year^2) + (delta.temp^2)), weather, 5000)
```
No zeroes for All Disasters against (Year + Temperature + (Year*Temperature) + (Year^2) + (Temerature^2))
Press - 592.9
Confidence - 95%

#Conclusions
For both the counts, Severe Storm and All Disaster, any combination of Year and Temperature seems to do the job. In this case, the lowest press becomes the decider of the function.
