---
title: "Assignment 3"
author: "Naman Shah"
date: "`r Sys.Date()`"
output: html_document
---

```{r, results='hide', include=FALSE, echo=TRUE}
library('moderndive')
library('tidyverse')
library('skimr')
source('fixregday1.pck')
source("regday1.pck")
```
## Let's build the models

```{r}
DroughtLogit.year <- glm(Drought.Count~Year, family = binomial (link = logit),
                          data = NOAAGISSWD)
DroughtLogit.temp <- glm(Drought.Count~delta.temp, family = binomial (link = logit),
                          data = NOAAGISSWD)
DroughtLogit.yearAndtemp <- glm(Drought.Count~Year + delta.temp, family = binomial (link = logit),
                                 data = NOAAGISSWD)
DroughtLogit.yearWithtemp <- glm(Drought.Count~Year * delta.temp, family = binomial (link = logit),
                                  data = NOAAGISSWD)

```
# Drought Compared to Year
```{r}
summary(DroughtLogit.year)
```
AIC: 51.153

Null deviance: 53.413 on 43 degrees

Residual deviance: 47.153 on 42 degrees

# Drought Compared to Temperature
```{r}
summary(DroughtLogit.temp)
```
AIC: 51.335

Null deviance: 53.413 on 43 degrees

Residual deviance: 47.335 on 42 degrees

# Drought Compared to Year+Temperature
```{r}
summary(DroughtLogit.yearAndtemp)
```
AIC: 52.995

Null deviance: 53.413 on 43 degrees

Residual deviance: 46.995 on 41 degrees

# Drought Compared to Year*Temperature
```{r}
summary(DroughtLogit.yearWithtemp)
```
AIC: 54.529

Null deviance: 53.413 on 43 degrees

Residual deviance: 46.529 on 40 degrees 

# Best Model for Drought
As we can see, the best model according to the lowest AIC is Drought Compared to Year. 

In the binomial functions using the logit link method, we see the AIC for the Drought Count compared to Year is 51.153, making it the most effective to predict Drought Count.